---
title: "TA Section: R Markdown, Github, Data exploration and GGplot"
subtitle: "QMSS Data Visualization - Spring 2021 - Week 02"
author: "Nico Rojas"
output:
  html_document:
    css: style.css
    toc: true
    self_contained: true
    keep_md: true
  ioslides_presentation:
    smaller: yes
    keep_md: true
  pdf_document:
    toc: true
    df_print: kable
urlcolor: red
font-family: Helvetica
autosize: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, message = F, warning = F}
library(gunsales) # NYtimes 
library(tidyverse) # Note: This encompasses dplyr, tidyr and ggplot
library(XML)
library(RCurl)
library(dplyr)
library(readr)
library(rvest)
library(stringr)
library(readxl)
library(knitr) # This one is for the table (kable)
library(kableExtra)
library(ggthemes) # This library is for making graphs prettier
library(scales) # This library is for fixing the scales in the end (dollar_sign)
```

```{r}
# This is a function to color text that I found on the R Markdown cookbok. 
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, 
      x)
  } else x
}
```

$~$

## 1. More resources for skills you will need. 

$~$

1. ***Basic R skills*** Check the Getting and transforming data in R section of the [syllabus](https://github.com/QMSS-G5063-2021/course_content). 
  - In particular, I strongly recommend mastering [section 5 of R for Data Science](https://r4ds.had.co.nz/transform.html) and the Data Camp introduction to R 
  - There is also an [unofficial solution manual written by Jeffrey Arnold](https://jrnold.github.io/r4ds-exercise-solutions/)
  - I also recommend downloading the slides from the course and also check [Yifei's TA repository](https://github.com/yifeitung/data_viz_review) (and his Github site in general, he has lots of insightful material there).
  
2. ***R Markdown skills *** Allows you to combine text and code and make your work reproducible. 
  - Some useful material are in this [course](https://rmarkdown.rstudio.com/lesson-1.html) by `R Studio`. The ***chunk section*** is incredibly useful. 
  - [R Makrdown cheat sheet](https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf) 
  - You can find more exhaustive material in  [2020 R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/) and [RMarkdown for Scientists](https://rmd4sci.njtierney.com/)
  
3. ***GitHub*** Check the [syllabus](https://github.com/QMSS-G5063-2021/course_content). 
  - For starters, [try this guide](https://guides.github.com/activities/hello-world/)

 $~$

```{r, include = F, eval = F}
#This chunk of text is a testimony that it is useful to have some r markdown skills. I could include a chunk that appears nothing but in Markdown. 

# Did you know that if you press esc while watching  week 02 slides, you can see a general map of all the slides?
```

$~$

## 2. Some Basic R Markdown skills 

$~$

### 2.1 - Knit and Chunks.  

* ***Knit function***: This function takes an input file, extracts the R code in it according to a list of patterns, evaluates the code and writes the output in another file.   

* ***Chunks***: A code chunk is a runable piece of code. It can be an R code, but also a Python or STATA code (or others). 
    - Chunks have options. All the options are [here](https://yihui.org/knitr/options/). 
    - There is an option called cache, that allows you to run long codes faster when you have already run them once.  
    - Three remarkably useful options:
    
$~$
    
    
| Cunk option      | Use | 
|-------------|---------------|
| echo  |	Determines whether to display the source code in the output document. |
| warning  |	Determines whether to preserve warnings |
| message  |	Determines whether to preserve messages |
 

#### Your turn {.exercise} 
- Generate a blank R Markdown and save it somewhere. 
- knit the document
- Write `library(tidyverse)` on one of the chunks.
- knit the document while checking what happens when echo = T or echo = F. 
- Do the same for "warning" and "message" in the chunk that has library(tidyverse). 
- You have 3 minutes. 

$~$

### 2.2 - Shortcuts. 

A short-list of useful short-cuts to work more efficiently (from [`RMarkdown for Scientists`](https://rmd4sci.njtierney.com/)). 

| Action      | Windows/Linux | Mac  | 
|-------------|---------------|----------------|
| Show Keyboard Shortcut Reference  |	Alt+Shift+K | Option+Shift+K |
| Knit document | Ctrl + Shift + K | Cmd + Shift + K |
| Insert Chunk  | Ctrl + Alt + I   | Cmd + Option + I |
| Run Current Chunk | Ctrl + Alt + C | Cmd + Option + C | 
| Jump to  	Shift+Alt+J 	| Cmd+Shift+Option+J |
| Create multiple cursors | Ctrl + Alt + Up/Down | option + control + Up/Down|
| Delete the current line | Ctrl + D | Cmd + D|
| Un/Comment out a line   | Ctrl + Shift + C | Cmd + Shift + C| 
| Reformat Section | Ctrl + Shift + A | Cmd + Shift + A| 

#### Your turn {.exercise} 
- Using the Keyboard Shortcut Reference, find the keyboard shortcut for inserting a pipe character ( %>% )
- Generate a blank R Markdown and spend 3 minutes using these commands. 

$~$

## 3. Github  

$~$

You need to be able to do, at least, the following actions on Github: 

| Action  | 
|-------------|
| 1) Create and use a repository  |	
| 2) Clone and use and existing repository|	
| 3) Start and manage a new branch  |
| 4) Make changes to a file, and push them to GitHub as commits |	
| 5) Open and merge a pull request |	

I strongly encourge you to do the excercise that is in the Hello World. [Git-hub guide](https://guides.github.com/activities/hello-world/)

Here we are going to ***clone a repository*** and ***make changes to a file and push them**** to the main branch. 

***Cloning a repository*** is creating a local copy of an existing repository on your computer. 

***Commmitting a change *** After modifying a file in the cloned repository you can commit a change. You are saving a change *locally* into a ***branch*** (either the main branch or other), with a name and a description of what this change makes.  

***Pushing *** updates your commits in to the remote repository. You can only push the branch to GitHub if you have write access to the repository. Nevertheless, you can always create a branch in GitHub Desktop if you have read access to a repository


#### Your turn {.exercise} 
- Clone the sandbox in your computer
- Generate a folder with your UNI and an empty text file inside.  
- Commit the change on your GitHub client 
- Can you push that change into the remote repository? Why? 
- Push the change if you can. 

$~$

## 4. (Very) Basic R Skills 

$~$

A minimum list of commands and things you need to master *now*.  

```{r, echo = F}
col1  <- c("Everything in the Introduction to R in Data Camp", "filter()", "select()", "arrange()", "mutate()", "group_by() and summarise()", "n()", "Everything about factors (Chapter 15)") 
col2 <- c("Check the outline of the data camp course and do whatever you don't know", "subset rows (observations)", "sorts rows", "keep or drop columns (observations, tip: use it as dplyr::select)", "generates a column", "group_by() gives the instruction to operate within subsets, summarise() collapses the data by using a function", "counts", "categorical variables with a fix set of values")
table <- data.frame(col1 = col1, col2 = col2)
```

```{r}
kable(table, col.names = c("Commands", "Description")) %>% kable_styling("striped") %>% row_spec(0, color = "red", font_size = 18)
```
$~$

## 5. R Skills: Data Exploration. 

$~$

Very quick review based on Exercise 2. Check the following table with commands. 

```{r, echo = F}
col1 <- c("str()","glimpse()","head()", "tail()","view()","just run the name of the object","names(), as.data.frame(names())", "summary()", "unique()", "table()")
col2 <- c("Check the structure of the data","Check the structure of the data (dplyr)", "Check the first few rows", "Check the last few rows", "View the data 1", "View the data 2", "Names of the variables",  "General description of all the variables***", "Dplyr command that can be used to check for duplicates (crucial for merging)", "Frequency table or two way table (use exclude = NULL for showing NA)") 
```

```{r}
kable(table, col.names = c("Commands", "Description")) %>% kable_styling("striped") %>% row_spec(0, color = "red", font_size = 18)
```

### Let's try some of the ones that have not been used yet in class. 

```{r, warning = F, message = F, echo = T}
# Data-source 1: Gunsales - NYTimes
gun_sales <- analysis()
# Note that if we hadn't use message = F warning = F, the final output would have been messier.
```

### Quick summary of all the data - NOTE: IT ALSO COUNTS MISSNIG DATA!

Several times the data provided during the courses comes from clean and nice data-sets. Don't forget to check whether some of your data may have substantial problems like missing data or unwanted duplicates that will ruin your analysis and data-merging.   

```{r,echo =  T}
summary(gun_sales)
```

### Quick List of all the variables in the data-set
```{r, message = F, warning = F, echo = T}
names(gun_sales)
as.data.frame(names(gun_sales))
```

#### Your turn {.exercise} 
On your own: Try all these commands on the gun_sales data_set. 

$~$

## 6. ggplot 

$~$

Run the following code to download an individual data-set for gun-related deaths in the US 

```{r, echo = T}
gun_inddeaths <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/guns-data/master/full_data.csv")
gun_inddeaths <- gun_inddeaths[,2:dim(gun_inddeaths)[2]]
```

Answer the following question: 

***What are the individual level patterns of gun deaths?***

$~$

### 6.1) What data do we have? 

```{r, echo = T, message = F, warning = F}
glimpse(gun_inddeaths) 
```

Let's say we are interested in exploring the relationship between race and intent. We can have a closer look to both variables.  

```{r, echo = T}
ftable(gun_inddeaths$intent, exclude = NULL) # One missing value 
ftable(gun_inddeaths$race, exclude = NULL)  # No missing values
table(gun_inddeaths$race, gun_inddeaths$intent , exclude = NULL)  
```

$~$

### 6.2) What are the ggplot alternatives available?  

Looks like a very open question. What is the more suitable geom available in `ggplot2`? 

Let's look at the [ggplot cheat sheet](https://rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf)   

```{r}
url <- "https://drive.google.com/uc?export=download&id=1iTbH8X8xlhiIiMXT8xPSeedfzQUZwecl"
```
![](`r url`)

Another way to check for the alternatives available is by visiting graph galleries. 

- Look to the ggplot section of the [R graph gallery](https://www.r-graph-gallery.com/ggplot2-package.html)
- Or [these top 50 ggplot](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html) visualizations bySelva Prabhakaran

Graph galleries also have codes that you can easily adapt to your purposes. 

$~$

### 6.3) Ok, we want to do a geom_count

From the ggplot cheat sheet we know that geom_count is an alternative to visualize two discrete variables. Let's try it: 

```{r, message = F, warning = F, echo = TRUE}
ggplot(gun_inddeaths, aes(intent, race)) + 
  geom_count()
```

It looks difficult to compare, specially because of the frequencies in each race.... 

What about the bar_plots from class with `fill` position. 

```{r, echo = TRUE}
ggplot(gun_inddeaths) + 
  geom_bar(aes(x = race, fill = intent), position = "fill") 
```

It looks, much better: easier to compare between races. 

#### We found a pattern we want to emphasize: {.exercise2} 
Blacks and Hispanics are less likely to die because of a suicide compare to the rest of the gun related-deaths. 

#### Let's list all the reasons why it still is different get the information we want from it:   
 
- There is NA category for missing values has very few values. (How do we know?)
- Helping the audience: 
  * Race ordering is alphabetical. We are not "making it easier for the audience" to compare between races.  
  * Intent ordering is alphabetical. We are not "making it easier for the audience" to compare between intents.  
- The legend "Intent", the title "race" and the title "count" seems unnecessary. 
- The labels in the bottom are unreadable
- Suicides and "less likely" are not emphasized compared to the other types of intents. 
- Can we make some tweaks to make the visualization more: "1) Trustful 2) Functional 3) Beautiful 4) Insightful"?

#### 

$~$

Let's fix all of this:

### 6.4) There is NA category for missing values has very few values.

```{r, echo = TRUE}
# How many NA?
ftable(gun_inddeaths$intent, exclude = NULL) # Only one!
ftable(gun_inddeaths$race, exclude = NULL)

# Let's fix this: 
ggplot(gun_inddeaths %>% drop_na(intent)) +  # We drop the only NA in intent 
  geom_bar(aes(x = race, fill = intent), position = "fill") 
```

$~$

### 6.5) Ordering: "Ordering improves visual perception"  

How can we work with discrete variables?: With factors.  

```{r, echo = TRUE}
gun_inddeaths$race <- factor(gun_inddeaths$race, levels = c("Black", "Hispanic", "Asian/Pacific Islander", "Native American/Native Alaskan", "White"))
gun_inddeaths$intent <- factor(gun_inddeaths$intent, levels = c("Undetermined", "Accidental", "Homicide", "Suicide"))

ggplot(gun_inddeaths %>% drop_na(intent)) +  # We drop the only NA in intent 
  geom_bar(aes(x = race, fill = intent), position = "fill") 
```
$~$

### 6.6) Drop redundant info (and now ilegible)

```{r, echo = TRUE}
ggplot(gun_inddeaths %>% drop_na(intent)) +  # We drop the only NA in intent 
  geom_bar(aes(x = factor(race, levels = c("Black", "Hispanic", "Asian/Pacific Islander", "Native American/Native Alaskan", "White")), fill = intent), position = "fill") + 
  xlab(NULL) + ylab(NULL) + theme(legend.title = element_blank())
```
$~$

### 6.7) The labels in the bottom are unreadable

* What if:
  - We shorten the labels. 
  - We move the legend to the bottom 

```{r, echo = TRUE}
levels(gun_inddeaths$race)[levels(gun_inddeaths$race)=="Asian/Pacific Islander"] <- "Asian/P. Islander"
levels(gun_inddeaths$race)[levels(gun_inddeaths$race)=="Native American/Native Alaskan"] <- "N. American/N. Alaskan"

ggplot(gun_inddeaths %>% drop_na(intent)) +  # We drop the only NA in intent 
  geom_bar(aes(x = factor(race, levels = c("Black", "Hispanic", "Asian/P. Islander", "N. American/N. Alaskan", "White")), fill = intent), position = "fill") + 
  xlab(NULL) + ylab(NULL) + theme(legend.title = element_blank()) +  theme(axis.text.x = element_text(size = 9)) +   theme(legend.title = element_blank(), legend.position =  "bottom")

```

$~$

### 6.8) Suicides and "less likely" are not emphasized compared to the of types of deaths. 

- Let's emphasize suicides by making a relative difference with the others types of deaths. 
  * Note that suicide is already emphasized in how we ordered the type of deaths. 
  * We want to use an additional channel to emphasize the information. 

- Let's emphasize "less" likely using an horizontal line at 50%. 

Beautiful? 

- Let's use a theme from ggthemes (look for this theme gallery [here](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/))

Functional?

- It's not that clear that the y axis label is percentages. Let's fix this.  


```{r, echo = TRUE, message = F}
levels(gun_inddeaths$race)[levels(gun_inddeaths$race)=="Asian/Pacific Islander"] <- "Asian/P. Islander"
levels(gun_inddeaths$race)[levels(gun_inddeaths$race)=="Native American/Native Alaskan"] <- "N. American/N. Alaskan"

ggplot(gun_inddeaths %>% drop_na(intent)) +  # We drop the only NA in intent 
  geom_bar(aes(x = factor(race, levels = c("Black", "Hispanic", "Asian/P. Islander", "N. American/N. Alaskan", "White")), fill = intent), position = "fill") + 
  theme(legend.title = element_blank()) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) + 
  geom_hline(yintercept=0.5, linetype="dashed", color = "black", size = 1) +  
  labs(x = NULL, y = NULL, 
       title = "Types of gun-related deaths by race \n", caption = "Source: FiveThirtyEight data.") + 
  scale_fill_manual("", values = c("#6666FF", "#99CCFF", "#CCCCFF", "red")) + theme_fivethirtyeight() + scale_y_continuous(labels = dollar_format(suffix = "%", prefix = "")) + scale_y_continuous(labels=percent_format()) + theme(axis.text.x = element_text(size = 9))
```

