---
title: "TA Session: Tools for data exploration and ggplot"
subtitle: "QMSS Data Visualization - Spring 2021"
author: "Nico Rojas"

widgets: []
mode: selfcontained
always_allow_html: yes
output:
  ioslides_presentation:
    incremental: false
    smaller: false
---

```{r packages, message = FALSE, warning = FALSE, echo=FALSE}
require(knitr)
require(leaflet)
require(shiny)
require(rgdal)
require(RColorBrewer)
```


```{r Setup, include=FALSE, results='hide', warning=FALSE}
## Normal Setup From Here
library(knitr)

opts_chunk$set(fig.path="images/",
               cache.path="cache/",
            #   dev=c("png","pdf"),
            #  fig.width=5,
            #  fig.height=4.5,
            #  dpi=300,
            #  fig.show="hold",
            #  fig.lp="fig:",
               cache=TRUE,
            #  par=TRUE,
               echo=TRUE,
               message=FALSE,
               warning=FALSE)
```

## Outline

- Tools that you need for this course 
- Tools for Data Exploration
- Example: Visualizing the World Development Indicators in Shiny
 
 
 
# Final Project

## Final Presentations - Details

- Each group has 10min time to present and for a short Q&A.
- Prepare to present for a few minutes. Better less, to allow for a bit of discussion.
- Remember: Everyone can look at your output even before class or with you.

## Feedback to other project

- Each student will provide comments about the visualizations used in another project. These comments can then be used to inform some final adjustments before submission.  
- We will post links to all final projects on a website.
- Everyone will receive an email specifying which project they should comment on.
- Comments are anonymous and shared on Github.


## How is the final project evaluated?

1. **Visualization**  
    - Is your visualization effective in answering the questions you intended to address? Is the storytelling clear?  
    - Do the visualizations follow visualization principles? 
    - Is it informative and interesting for a user? 
    - Does it allow some sensible interactive exploration?  
2. **Implementation**  
    - What is the quality of your implementation? 
    - Is it appropriately polished, robust, and reliable?  
3. **Presentation**     
    - Are your web site and presentation clear, engaging, and effective?
    
