---
title: "TA Section: Plotly"
subtitle: "QMSS Data Visualization - Spring 2021 - Week 04"
author:  "Nico Rojas - nr2670@tc.columbia.edu"
output:
  html_document:
    css: style.css
    toc: true
    self_contained: true
    keep_md: true
  ioslides_presentation:
    smaller: yes
    keep_md: true
  pdf_document:
    toc: true
    df_print: kable
urlcolor: red
font-family: Helvetica
autosize: yes
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, message = F, warning = F}
library(tidyverse)
library(plotly) 
library(babynames)
library(NutrienTrackeR) #Package with nutrition data-set
```

$~$

## 0. Plotly resources. 

$~$

Some of the resources I find more useful to learn plotly using R:

1. @sievert_interactive_2020 - [plotly for R](https://plotly-r.com/index.html) - A full reference website for using plotly in R. (from the Syllabus)
2. Paul C. Bauer & Richard Traunmüller (2016) - [Interactive Data Visualization](https://bookdown.org/paulcbauer/idv2/) -  Book with all the material from a Workshop Data Visualization 
3. The [Plotly R](https://plotly.com/r/) website  
 
This session will rely on this material. 

$~$

## 1. Why ggplotly alone is not enough for implementing interactivity? 

$~$

It's tempting to think that ggploty will cover all the interactivity we might need and will save us the time of learning plotly coding. However, there are at least four reasons 

1) Not all ggplot objects can be converted to plotly objects 
2) Not all the parts of a ggplot object can be converted to plotly objects. 
3) Adjusting graphs that come from ggplotly can be cumbersome 
4) ggplot is not interactive, and therefore some interactive options are missing

## 1.1 ggplotly in combination with plotly syntax 

With respect to 4), we can use an example based on @sievert_interactive_2020 (Ch.33) to show some interactivity that cannot be displayed using ggplotly alone.  

This is an interactive "babynames" graph from week 01 exercise built based on ggplotly. 

```{r,echo = T}
p <- ggplot(babynames, aes(year, n)) + 
  geom_line(data = filter(babynames, name=="Alex"), aes(color = sex)) + ggtitle("ggplotly babynames")
ggplotly(p) 
```

Note that when you click into one point of the curve you only get data from data of the part of the curve.

***what if you wanted to know what's happening for all the curves for any given year?*** 

That option is not part of ggplot, because ggplot is not interactive. 

You can add plotly code after using ggplotly. We can change the layout to include the desired interactivity with `layout(hovermode = "x")`

```{r}
ggplotly(p  + ggtitle("ggplotly babynames adding hovermode")) %>%
  layout(hovermode = "x")
```
Other cool alternatives that are only in plotly are including a tool for zooming for a certain range of X's (`rangeslider()`). We can also use the `dynamicTicks = TRUE` option to adjust the xaxis while zooming.  

```{r, echo = T}
ggplotly(p  + ggtitle("ggplotly babynames adding hovermode and rangeslider")
         , dynamicTicks = TRUE) %>% 
  rangeslider() %>%
  layout(hovermode = "x")
```
$~$

## 2. Writing a plotly code   

$~$

### 2.1 - General aspects of plotly coding  

Every aspect of a plotly chart is determined in one of three aspects:  plot_ly(), add_trace() and layout()

In plotly a ***trace*** defines a mapping from data and visuals. Every trace has a type (e.g., histogram, pie, scatter, etc). plot_ly() alone already plots a "trace".  

* If you don't specify a layout(), plotly adds a default layout. 
* Even if you don't specify a type it will plot a default, unlike ggplot.

```{r, warning = F, message = F, echo = T}
plot_ly(economics, x = ~date, y = ~uempmed) # We didn't specified the type! Default: Scatterplot 
```

Also, unlike ggplot, you can use the pipe ( %>% ) and dplyr syntax with plotly coding. For example: 

```{r, echo = T, eval = F}
#Instead of this:
plot_ly(economics, x = ~date, y = ~uempmed)

#You can do this: 
economics %>% 
  plot_ly(x = ~date, y = ~uempmed)
```

Also, you can do group_by operations in a syntax: 

```{r, eval = F, echo = T, warning = F, message = F}
diamonds %>%
  group_by(depth) %>% 
  summarise(price = mean(price)) %>%
  ungroup() %>% 
  plot_ly(x = ~depth, y = ~price)
```

One advantage of the integration with dplyr and the pipe is that it allows you avoid using the plotly syntax with several parenthesis.  

Even though just using plotly() already plots a "trace" you can add another one with add_trace(). 

Note, that instead of using add_trace(), there is a family of add functions: add_histogram(), add_bars(), etc .... 

```{r, eval = T}
set.seed( 123 )
x = 1 : 100
y1 = 2*x + rnorm ( 100 )
y2 = - 2*x + rnorm ( 100 )
```

Look at this graph:

```{r, echo = T, warning = F, message = F}
plot_ly( y = y1, x = x , type = "scatter" ) %>%
add_bars( y = y2 )
```

It's the same as this graph: 

```{r, echo = T, warning = F, message = F}
plot_ly( x = x , type = "scatter" ) %>%
add_trace( y = y2, type = "bar")
```

***layout()*** apply to the rest of the graph, like the title, axis settings, or annotations

```{r, eval = F, include = F}
plot_ly(economics, x = ~date, y = ~uempmed) %>% 
  layout(title = "My beatiful graph")
```

### 2.2 - Plotly attributes  

***What are all the possible attributes that can be changed?** 

Let's take a look to the following syntax from the plotly website: 

```{r, echo = T, include = T, eval = T, message = F, warning = F}
# Note: Add: https://plotly.com before most of the links below. 

plot_ly(economics,
         type = "scatter",        # all "scatter" attributes: https://plotly.com/r/reference/#scatter
         x = ~date,               # more about scatter's "x": /r/reference/#scatter-x
         y = ~uempmed,            # more about scatter's "y": /r/reference/#scatter-y
         name = "unemployment",   # more about scatter's "name": /r/reference/#scatter-name
         marker = list(           # marker is a named list, valid keys: /r/reference/#scatter-marker
           color="#862626"        # more about marker's "color" attribute: /r/reference/#scatter-marker-color
         )) %>% 
  add_trace(x = ~date,                                         # scatter's "x": /r/reference/#scatter-x
        y = ~fitted((loess(uempmed ~ as.numeric(date)))),  # scatter's "y": /r/reference/#scatter-y
        mode = "lines",                                    # scatter's "y": /r/reference/#scatter-mode
        line = list(                                       # line is a named list, valid keys: /r/reference/#scatter-line
          color = "#264E86",                               # line's "color": /r/reference/#scatter-line-color
          dash = "dashed"                                  # line's "dash" property: /r/reference/#scatter-line-dash
        )
) %>% 
  layout(                        # all of layout's properties: /r/reference/#layout
     title = list(text = "Unemployment", # layout's title: /r/reference/#layout-title
                  font = list(
                  size =  20)),
                  margin = list(autoexpand = T, 
                                t = 50), # Adds some margin to make the title visible
     xaxis = list(           # layout's xaxis is a named list. List of valid keys: /r/reference/#layout-xaxis
        title = "Time",      # xaxis's title: /r/reference/#layout-xaxis-title
        showgrid = F),       # xaxis's showgrid: /r/reference/#layout-xaxis-showgrid
     yaxis = list(           # layout's yaxis is a named list. List of valid keys: /r/reference/#layout-yaxis
        title = "uidx")     # yaxis's title: /r/reference/#layout-yaxis-title
)
```

Where are all the possible options? Sieveret recommends using the schema function over looking elsewhere. 

```{r, echo = T}
plotly::schema()
```
```{r}
url <- "https://drive.google.com/uc?export=download&id=1ssWxnYC4yxMOm-jjtFsRQwLYapeAwsf1"
```
![](`r url`)


#### Your turn {.exercise} 
* Copy the syntax from the scatter plot and modify the different attributes. 
* run schema() in a chunk and look for other attributes that can be modified. Recommendation: Start with the title.
* Change the type of plot, and look to the different attirbutes available on schema()  
* Find the "type" of trace that is added to the plot. Is it line?  
* Find the `hovermode` attirbute used in the first part of the section, and find out other attributes that could be used to combine ggplotly with plotly syntax.

$~$

## 3. An example: Horizontal bar chart    

$~$

In this example we will use a nutrition data-set from the USDA that can be loaded using the `NutrienTrackeR` package. We will graph the top 10 categories of in terms of fat in the United States. 

```{r, echo = T, warning = F, message =  F}
top10 <- data.frame(food_composition_data$USDA, stringsAsFactors = F) %>% 
  group_by(food_group) %>%  
  summarize(mean_fat = mean(as.numeric(Total.lipid..fat...g.), na.rm = T) ) %>% 
  ungroup() %>% 
  arrange(desc(mean_fat)) %>% 
  filter(row_number()<11) %>% 
  mutate(mean_fat = round(mean_fat, digits = 1))
```

```{r, echo = T}
fig1 <- plot_ly(top10, x =  ~mean_fat, y = ~reorder(food_group, mean_fat),
              type = 'bar', orientation = 'h',
              marker = list(color = "#fc03df", #Google HEX color picker
                            line = list(color = 'rgba(50, 171, 96, 1.0)', width = 1))) 

fig1 <- fig1 %>% layout(title = "Top 10 categories of food with more fat in the US\n", 
                        yaxis = list(title = FALSE, domain= c(10, 67)), 
                        xaxis = list(title = "grams per 100 grams of food"), 
                        margin = list( autoexpand = TRUE, t = 80, b = 90), # put some margin on top so the title can be seen  
                        annotations =   list(x = 1, y = - 0.3, #position of text adjust as needed 
                text = "<i>Source: USDA Standard reference dataset</i>", # Note the use of HTML language to add italics 
                showarrow = F, xref='paper', yref='paper', 
                xanchor='right', yanchor='auto', xshift=0, yshift=0,
                font=list(size=12)))
fig1
```

$~$

## 4. Short survey    

$~$

[Here](https://forms.gle/WPSZBZ5CqvmV4RdT8) a short survey for asynchronous learners. It's anonymous, I just want to check how many people are looking this material and if you have any comment. Thank you!

***

## References


